---
title: "Analyzing Real Data with LRMoE"
description: >
  An example of LRMoE on fitting real claim frequency data.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analyzing Real Data with LRMoE}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
install.packages("xts") 
install.packages("sp")
install.packages("CASdatasets", repos = "http://cas.uqam.ca/pub/", type="source")
library(LRMoE)
library(reshape2)
library(matrixStats)
library(CASdatasets)
glm_fit = readRDS("realdata_glm_model.rda")
df = readRDS("realdata_converted_data.rda")
init = readRDS("realdata_init_3comp.rda")
LRMoE_fit = readRDS("realdata_LRMoE_3Poisson.rda")
# data("LRMoEDemoData")
# fitted_model = readRDS("fitted_model_true.rda")
# fitted_model_new = readRDS("fitted_model_new-raw.rda")
```

## Introduction

In this document, we will analyze a real dataset of claim frequency with LRMoE.
We will compare how it performs compared with the classical Poisson generalized
linear model (GLM).

## Overview of Dataset

We will use the `ausprivauto0405` dataset available in the R package [`CASdatasets`](http://cas.uqam.ca/).

```{r}
library(CASdatasets)
data(ausprivauto0405)
```

The dataset contains some typical covariates used in predicting auto-insurance claim frequency and severity.
For simplicity, we will only consider the problem of modelling claim frequency. A brief summary of the dataset
is given below.

```{r}
summary(ausprivauto0405)
```



## Poisson GLM

As a benchmark, we will first fit a Poisson GLM (with `log` link) to claim frequency. Note that `Exposure` represents how long the policy has been in force / observed, which is used as an `offset` term in model fitting.
The usual model summary is also given below.

```{r, eval=FALSE}
glm_fit = glm(ClaimNb ~ Gender + DrivAge +
            VehBody + VehAge + VehValue,
            offset = Exposure,
            family = poisson(link = "log"),
            data = ausprivauto0405)
```

```{r}
summary(glm_fit)
```

## Data Transformation

In order to fit an LRMoE, we first need to transform the dataset. Since `DriverAge`, `VehBody` and
`VehAge` are factors, we will use `model.matrix` to augment a dummy matrix of zeros and ones to represent them.

```{r, eval=FALSE}
# 1: Intercept
df = data.frame(Intercept = rep(1, nrow(ausprivauto0405)))
# 2: Gender
df = cbind(df, GenderMale = model.matrix(~Gender, data = ausprivauto0405)[,2])
# 3-7: DrivAge
df = cbind(df, model.matrix(~DrivAge, data = ausprivauto0405)[,2:6])
# 8-19: VehBody
df = cbind(df, model.matrix(~VehBody, data = ausprivauto0405)[,2:13])
# 20-22: VehAge
df = cbind(df, VehAge = model.matrix(~VehAge, data = ausprivauto0405)[,2:4])
# 23: VehValue
df =  cbind(df, VehValue = ausprivauto0405$VehValue)


# 24: Exposure
df = cbind(df, Exposure = ausprivauto0405$Exposure)


# 25: ClaimNb
df = cbind(df, ClaimNb = ausprivauto0405$ClaimNb)

df = as.matrix(df)
```


## Model Initialization

When fitting an LRMoE model, a good starting point usually reduces run time and can potentially land on a better fit. In the **LRMoE** package, we provide a function for parameter initialization based on the Clustered Method-of-Moments (CMM).

The following demonstrates how to initialize a 3-component LRMoE model.

```{r, eval=TRUE}
set.seed(2021)
X = as.matrix(df[,1:23])
exposure = as.matrix(df[,24])
Y = as.matrix(df[,25])
```

```{r, eval=FALSE}
init = LRMoE::cmm_init(Y = Y, X = X, n_comp = 3, 
                       type = c("discrete"), exact_Y = TRUE)
```

The `init` object contains a list of summary statistics of `Y`, as well as a set of parameter
initialization. Let us consider a 3-component Poisson mixture.

```{r, eval=TRUE}
# Initialization of alpha
alpha_init = init$alpha_init
alpha_init
```

```{r, eval=TRUE}
# Initialization of component distributions
# The list indices are specified as:
# [[1]]: the 1st dimension of the response of Y
# [[1]]/[[2]]/[[3]]: the 1st, 2nd and 3rd latent component/group
# [[5]]: the 5th ponential choice of expert is zipoisson
comp_dist = matrix(list(init$params_init[[1]][[1]][[1]]$distribution,
                        init$params_init[[1]][[2]][[1]]$distribution,
                        init$params_init[[1]][[3]][[1]]$distribution),
                   nrow = 1, byrow = TRUE)
params_init = matrix(list(init$params_init[[1]][[1]][[1]]$params,
                          init$params_init[[1]][[2]][[1]]$params,
                          init$params_init[[1]][[3]][[1]]$params),
                     nrow = 1, byrow = TRUE)
```

If interested, we can also check some summary statistics of `Y` based on the initialization.
For example, the proportion of zeros and the mean of positive `Y`'s are given below. These summary statistics may help with selecting the appropriate expert functions for each latent component/group.
```{r}
init$zero_y
```

```{r}
init$mean_y_pos
```

## Poisson LRMoE

With an initialization of model, we can now fit the LRMoE with zero-inflated Poisson experts.
Note that `exposure` is also taken into account. In the case of Poisson distribution with parameter $\lambda$, a policy with exposure $E_i$ would have claim frequency distribution $Poisson(\lambda E_i)$. The same rate parameter $\lambda$ is shared by all policyholders who have potentially different policy exposures.

(Note: **The fitting function takes more than 30 minutes to run.** Adjusting parameters such as a larger `eps` or smaller `ecm_iter_max` will provide a quicker, but potentially worse, model fit.)

```{r, eval=FALSE}
LRMoE_fit = LRMoE::FitLRMoE(Y = Y, X = X,
                            alpha_init = alpha_init,
                            comp_dist = comp_dist, params_list = params_init,
                            exposure = exposure,
                            exact_Y = TRUE,
                            eps = 0.05, ecm_iter_max = 25)
```

We can examine the fitted model as follows, as well as its loglikelihood, AIC and BIC.
Note that, although we are fitting the same dataset, the result above is different from that presented in the Actuary Magazine. This is most likely due to a different initialization as well as stopping conditions. In practice, such differences are unlikely to have a material impact, as long as the overall goodness-of-fit is similar for these different models.

```{r}
LRMoE_fit$alpha_fit
```

```{r}
LRMoE::print_expert_matrix(LRMoE_fit$model_fit)
```

```{r}
LRMoE_fit$ll
```

```{r}
LRMoE_fit$AIC
```

```{r}
LRMoE_fit$BIC
```


